---
title: "Instacart Analysis with Flexdashboard"
author: "Zhezheng Jin"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    navbar:
      - { title: "Home", href: index.html }
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(plotly)
library(p8105.datasets)
library(flexdashboard)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Let's take a look at the dataset 

```{r}
instacart
```

Data wrangling

```{r}
instacart <- instacart %>%
  janitor::clean_names() %>%
  mutate(order_dow = case_when(
    order_dow == 0 ~ "Sun",
    order_dow == 1 ~ "Mon",
    order_dow == 2 ~ "Tue",
    order_dow == 3 ~ "Wed",
    order_dow == 4 ~ "Thu",
    order_dow == 5 ~ "Fri",
    order_dow == 6 ~ "Sat",
    TRUE ~ as.character(order_dow)
  ),
  order_dow = factor(order_dow, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")))
```


From the dataset, we know there are `r ncol(instacart)` variables in this dataset: `r paste(colnames(instacart), collapse=", ")`. With these information, I am going to create 3 distinct plots:

1. A histogram showing the number of orders for each day of the week.

2. A line plot showing the average number of Orders by hour within each day of the week.

3. A box plot of distribution of Add to Cart Order for top 5 Products.


## Histogram

```{r}
plot_ly(instacart, x = ~order_dow) %>%
  add_histogram(name = "Count of Orders") %>%
  layout(title = "Distribution of Orders by Day of the Week",
         xaxis = list(title = 'Day of the Week'), 
         yaxis = list(title = 'Count of Orders'))
```

## Line plot

```{r}
instacart %>%
  group_by(order_dow, order_hour_of_day) %>%
  summarise(avg_orders = n() / n_distinct(user_id)) %>%
  arrange(order_dow, order_hour_of_day) %>%
  plot_ly(x = ~order_hour_of_day, 
          y = ~avg_orders, 
          color = ~order_dow,
          type = "scatter", 
          mode = "lines+markers", 
          marker = list(size = 7),
          line = list(width = 2)) %>%
  layout(title = "Average Number of Orders by Hour within each Day of the Week",
         xaxis = list(title = "Hour of the Day"),
         yaxis = list(title = "Average Number of Orders"),
         legend = list(title = list(text = 'Day of the Week')),
         showlegend = TRUE)
```

## Box plot

```{r}
top_products <- instacart %>%
  group_by(product_name) %>%
  tally() %>%
  top_n(5, wt = n) %>%
  pull(product_name)

subset_data <- instacart %>%
  filter(product_name %in% top_products)

plot_ly(data = subset_data, 
        y = ~add_to_cart_order, 
        x = ~product_name, 
        type = "box",
        orientation = "v") %>%
  layout(title = "Distribution of Add to Cart Order for Top 5 Products",
         yaxis = list(title = "Add to Cart Order"),
         xaxis = list(title = "Product Name", tickangle = 15))
```



